Optimised Cost Considering Huffman Code for Biological Data Compression

### 2015-02-03 Article/Article.tex
finished

\documentclass[preprint,12pt]{elsarticle}

@@

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{lscape}
\usepackage[noend]{algpseudocode}
\usepackage{algcompatible}
\usepackage{multirow}

@@

%\author{}
\author[a,b]{Youcef Gheraibia\corref{cor1}}
\ead{Y.Gheraibia@hull.ac.uk}
\author[b]{Sohag Kabir\corref{cor2}}
\ead{s.kabir@hull.ac.uk}

\address{}

\address[a]{Department of Mathematics and Computer Science, University of Souk Ahras, Souk Ahras 41000, Algeria}

\address[b]{Department of Computer Science, University of Hull, Hull HU6 7RX, UK}

@@

\begin{keyword}
Data Compression, Huffman code, Cost Considering, Genetic algorithm, Biological data.
%% keywords here, in the form: keyword \sep keyword

@@

\section{Background}
\label{sec2}
\subsection{Issues on Biological Data Transmission}
The size of biological data including DNA sequences increase with an ever expanding rate and will be bigger and bigger in the future. These biological data are stored in biology databases. The exponential growth of these databases become a big problem to all biological data processing methods~\cite{Doug08}.
Different operations are applied to these data such as searching ~\cite{val10},e-mail attachment~\cite{chr09}, alignment ~\cite{che03}, and transmission on distributed computing platform~\cite{cha14}. Interestingly, biological data compression can play a key role in all sort of biological data processing.

@@

Applications of Huffman code are pervasive throughout
computer science. The algorithm to completely perform Huffman encoding and decoding is explained in \cite{Amst86}. It can be used effectively where there is a need for a compact code to represent a long series of a relatively small number of distinct bytes. For example, Table \ref{table1} shows 8 different ASCII characters, their frequencies, ASCII codes and the codewords generated for those symbols using Huffman code. It is seen from the table that the codeword to represent each character is compressed and the most frequent character gets the shortest code. In this example, the compression ratio obtained by Huffman code is 64.16\%.

@@

\section{Approach}
\label{sec3}
\subsection{Proposed Scheme}

@@

\subsection{Cost Considering Huffman code}
The classical Huffman algorithm aims at reducing total number of bits and it constructs a tree in a bottom up fashion. It is shown in \cite{golin98} that if the costs of letters are considered unequal then the straightforward bottom up greedy approach does not work. Authors in \cite{Kab14} uses a top down approach to build a binary tree considering unequal letter cost of bits. They considered cost (length) of 0 and 1 as integer constants $\alpha$ and $\beta$,   $\alpha < \beta$. Using the analogy of Morse code's `$\cdot$' and `$-$', the value of $\alpha$ and $\beta$ is set as 1 and 3 respectively. The complete algorithm to obtain an optimal prefix-free code for unequal letter cost is shown below. The input to the algorithm are the distinct triplets contained in the genome sequence to be encoded and their frequencies. The process of creating the binary tree starts with a single node (root node) and it is initialised with cost 0.  After that, two child (leaf) nodes are created for the root node, i.e., level of the tree is increased by one. Cost of the left child is calculated as the summation of the cost of its parent node and the length of the left arc, and cost of the right child is calculated as the summation of the cost of its parent node and the length of the right arc. Length of left and right arcs are actually the cost (length) of 0 ($\alpha$) and 1 ($\beta$) respectively. The next step is to take a child node with least cost and create two child nodes for it and make it a parent node. In this way, in every iteration the child node with least cost becomes a parent node with two new child nodes. Creation of new child nodes is stopped when total number of child nodes become equal to the number of distinct triplets needed to be encoded.

Now the tree $T$ is constructed and the cost of the tree actually depends on how the frequencies are assigned to the leaf nodes. The overall cost will be minimised if the leaves with highest cost always have smaller or equal weight (frequency). To fulfil this condition the leaves of the $T$ are enumerated in non-decreasing order of their cost, i.e., $cost(l_1)\leq cost(l_2)\leq \ldots \leq cost(l_n),$ and that $f_1 \geq f_2 \geq \ldots\geq f_n$, where $l_i$ and $f_i$ are leaf nodes and frequencies of distinct triplets respectively $for~ i=1,2,\ldots n$ . The frequency or weight of parent nodes are calculated as the sum of the frequencies of its child nodes, and it continues upwards until the root node is reached. After that, the algorithm checks for any possible conflicts between all pair of nodes. Two nodes are considered as conflicted if the node with a higher cost has higher frequency violating the above condition, i.e., if $cost(l_i)> cost(l_j)~and~f(l_i)>f(l_j)$, then there remains a conflict.  If there remains a conflict between nodes, then it is resolved by swapping the nodes and recalculating the cost of the tree downward and frequency of the nodes upward.  When all the conflicts if existed are resolved then the algorithm generates codes for each of the distinct triplets.   
\begin{algorithm}[!btph]
\caption{Cost-considering / Unequal bit cost Coding}
\label{alg1}

@@

\FOR{each distinct symbol $i$} 
\STATE $ Enqueue\left(max\_Q , frequency~[~i~]\right)$
\ENDFOR
\STATE create a root node
\STATE $cost~[~root~]\leftarrow 0$
\STATE $Enqueue\left(min\_Q , cost~[~root~]\right)$
\STATE Define costs of the left and right child of the binary tree
\REPEAT 
\STATE $cost\_of\_parent\_node\leftarrow Dequeue\left(min\_Q\right)$
\STATE create $left$ and $right$ child for this node
\STATE $cost~[~left\_child~]\leftarrow cost\_of\_parent\_node+left\_child\_cost$
\STATE $Enqueue\left(min\_Q , cost~[~left\_child~]\right)$
\STATE $cost~[~right\_child~]\leftarrow cost\_of\_parent\_node+right\_child\_cost$
\STATE $Enqueue\left(min\_Q , cost~[~right\_child~]\right)$
\STATE Mark parent node as explored
\UNTIL{$2\left(n-1\right)$ nodes are created}
\WHILE{$min\_Q\neq\emptyset$} 
\STATE $leaf\_node\leftarrow Dequeue\left(min\_Q\right)$
\STATE $frequency[leaf\_node]\leftarrow Dequeue\left(max\_Q\right)$
\ENDWHILE
\FOR{each parent node $j$} 
\STATE $frequency~[~j~]\leftarrow frequency~[~left\_child~]+frequency~[~right\_child~]$
\ENDFOR
\REPEAT
\IF{conflict between nodes} 
\STATE resolve conflict by swapping conflicted nodes
\STATE calculate and reassign cost of all affected nodes
\STATE calculate and reassign frequency of all affected nodes 
\ENDIF
\UNTIL{all conflicts are resolved }
%\algstore{myalg}
\end{algorithmic}
\end{algorithm}
\subsection{Optimisation of the Codes}
\subsubsection{Problem formulation}
The problem of finding the best allocation of codes to each triplet can be   modelled as an assignment problems, the problem is formulated as follows :\\\\
\textbf{Definition:} Given a set of codes $C=\left\{C_{1},C_{2}...C_{n}\right\}$, and a set of frequencies $F=\left\{F_{1},F_{2}...F_{n}\right\}$. For each code we have the length of the code $|C_{i}|$ (number of bits) and the cost of the code $S_{C_{i}}$, the objective is to assign to each frequency a code in order to minimise the total number of bits, while respecting the initially assigned total cost $S_{t}$ with a given penalty $\lambda\rightarrow [0,1]$. This penalty coefficient represents the allowed amount of cost that can be sacrificed to optimise the total number of bits.    
\\
The Objective Function is to:

@@

\subsubsection{Basic Genetic Algorithm}
Genetic Algorithm (GA) is a bio-inspired meta-heuristic algorithm developed by \cite{gen1}. GA is a stochastic optimisation algorithm imitate the natural evolution process of genomes. GA started by generating an initial population of random feasible solutions. The optimisation process of GA takes the initial population and generates a new population based on it.  The process can be described as follow:\\ 
Firstly, select two or more solutions from the current population by using one of the well-known selection techniques \cite{bli95}. These selected solutions will be considered as  parents. The second operation of the genetic algorithm is the crossover, which takes these parents as inputs to generate other new solutions considered as children. The third operation of the genetic algorithm is the mutation which ensures a good diversification in the search process. The new solutions can be mutated according to a given mutation probability. The mutation operation changes the value of one or more positions in the solution. The quality of each solution is verified by the fitness function which controls the evolution of the GA population by deletion of the worst solutions and insertion of the good solutions among  parents and sons. This processes is repeated until the stopped criteria is reached which can be the number of generation or if the population is stabilised.
\subsubsection{Optimised cost considering algorithm}
The main objective of the GA optimisation algorithm for bits minimisation OCCA is to assign each frequency a specific code. \textbf{In GA optimisation algorithm, the encoding of the solution into a chromosome is one of the most important issue to obtaining a good optimisation results}. The OCCA uses two fixed length arrays of size ``64" which is the number of combination for all nucleotides. The first array contains the frequencies of each triplets and the second array contains the cost of the codewords assigned to each triplets. In this way, our genetic algorithm will work with the two arrays and uses the entry index on the allocation process.

The genetic algorithm\textbf{ is a }stochastic algorithm based \textbf{on random evolution process}. \textbf{Generally the initial population is generated in a random affectation of codes to different triplets}. \textbf{In the OCCA algorithm, the population contain firstly the affectation given by the CCA. Secondly the rest of the population are randomly generated, all these generated solutions must satisfy the initial cost constraints, which is the cost of the CCA solution previously add to the population (step 1)}.

The evolution of the population is the key of the Genetic optimisation algorithm. During each generation the process start with the \textbf{selection of a part (set of solutions) of the population} to bread a new generation. In the literature many selection methods have been proposed to guide the population evolution \cite{bli95}. selection of a proportion (set of solutions) from the population  methods are varied from a random selection to heuristic based selection. We have chosen to select randomly the part of the population to be processed, as the heuristic methods are very time-consuming (step 3).
After that the operations of genetic algorithms are applied on the initial population to generate a new generation of the population (see figure 2). Firstly the crossover operation are applied to these two selected solution (considered as parents) to generate two new solutions (considered as sons)(step 4). In the literature many crossover techniques have been used in genetic algorithm \cite{osa14}, such as one-point crossover which divide the chromosome into two fragments and recombine the second fragments with the other chromosome's second fragment, two-point crossover which divide the chromosome into three fragments and recombine the middle fragment with  the middle fragments of the other chromosome. There are many other crossover techniques to allow a good convergence of the algorithm. In our case, we have used the two point crossover with two parameters, the first parameter $p_{c}$ is a random value in [0,63], which represent the first cut point and $p_{m}$ is a random value too in [0,63], which represent the number of position to be crossed. We used two random parameters to ensure a good diversification on the whole search space (step 4).
These two new generated children may contain conflict like finding a duplicated code allocated to two different frequencies in the solution, so a regulation step is done to ensure the correctness of the solution (see figure 2).
 Secondly these two new solutions are mutated according to a predefined probability $\gamma$, the best value of mutation rate is very problem specific (step 5). The value of $\gamma$ is fixed to $0.2$ to explorer a few position on the solution. The mutation operator used to maintain genetic diversity from one generation of a population of genetic algorithm chromosomes to the next. In our case, we have used the mutation as a random swap mutation operator (see figure 3).  Each newly generated solution must satisfy the cost constraint, these children must be valid with satisfying the cost constraint. The next step is to add these two new solutions (children) to the population (step 7) (see figure 3). Finally the new population are ranked by fitness (step 8), and the worst solutions are deleted until the initial size of the population are achieved  (step 9). The whole processes are repeated until the maximum number of operation is achieved (step 10). 
\begin{algorithm}[t]
\caption{Optimised cost considering algorithm}
\begin{algorithmic}[1]
\State Population initialization (P).
\REPEAT  
\State Select two solutions $S_{1},S_{2}$ form P.
\State Crossover $S_{1},S_{2}$ to generate $S_{11},S_{21}$ (Children).
\State Mutate $S_{11},S_{21}$.
\State Validate children with cost constraint (equation 4). 
\State Add children to population
\State Rank the population by fitness
\State Remove worst candidates until population limit
\UNTIL{Max number of generation not achieved} 
\EndWhile
\State Display the best solution from the population P;
\end{algorithmic}
\end{algorithm}

@@

\begin{figure}[!thpb]
\begin{center}
\includegraphics[scale=0.7]{Images/Drawing1.jpg}
\caption{Population Update for genetic algorithm}
\end{center}
\label{Fig2}
\end{figure}

\section{Results And Discussion}
\label{sec4}
The effectiveness of the approach has been evaluated with different real genomic biological data, these genomes were downloaded from a recent version of The National Center for Biotechnology Information (NCBI) available on $(http://www.ncbi.nlm.nih.gov)$ \cite{pruitt2009ncbi}. We focused on the sequences alone, ignoring any header and any other exogenous information. In table~\ref{table3} the different data sets are described with the size of each of them and the references on the biological data bank.
\begin{table}[!thpb]
\small
\label{table3}
\caption{Dataset description}
%\begin{center}
\centering
\begin{tabular}{c  c  c c}
\toprule
$\textbf{Data sets}$ &$\textbf{Name}$ &	$\textbf{Size (MB)}$ &	$\textbf{Reference}$ \\\hline
Genome 1& Mycobacterium smegmatis &  6.66 & CP009496  \\\hline

@@

\multirow{2}{*}{Genome 16}& Danio rerio & \multirow{2}{*}{56.14} &\multirow{2}{*}{CM002885}\\ & chromosome 1 &\\
\hline
\multirow{2}{*}{Genome 17}&Homo sapiens & \multirow{2}{*}{76.64 } &\multirow{2}{*}{  CM000680  }\\    & chromosome 18 &\\
\hline
\multirow{2}{*}{Genome 18}&Homo sapiens & \multirow{2}{*}{99.94} &\multirow{2}{*}{CM000684   }\\ & chromosome 22&\\
\hline
\bottomrule
\end{tabular}
%\end{center}
\end{table}

\begin{table}[h]
\small
\label{table4}
\caption{OCCA comparison for cost and size with different approaches with different datasets}
%\begin{center}
\centering
\begin{tabular}{c | c c | c c | c c}
\toprule
 & \multicolumn{2}{c}{Huffman Algorithm} & \multicolumn{2}{c}{CCA }& \multicolumn{2}{c}{OCCA}\\\hline
$\textbf{Data sets}$ & $\textbf{Cost}$	& $\textbf{Size(MB)}$ &	$\textbf{Cost}$&	$\textbf{Size(MB)}$&$\textbf{Cost}$&	$\textbf{Size(MB)}$
\\\hline
Genome 1& 76787151  &	4.44	&67416213   & 4.92	& 67416213 & 4.82 \\\hline
Genome 2& 100425402 &  	5.81	& 88430665	& 6.51	& 88430665&  6.35 \\\hline
Genome 3& 75940155  &	4.39    &66745619	& 4.87&	66745619 & 4.79 \\\hline
Genome 4& 103552729 &	5.96    &90821835&	6.69 &	90821835  & 6.47\\\hline
Genome 5& 82234926  &	4.74	&71963876	& 5.27 &	71963876 &5.17 \\\hline
Genome 6& 83454842  &   4.81	&73038795	& 5.34	&73038795 & 5.24\\\hline
Genome 7& 92539488  &	5.36    &81416359   &5.93	&81416359 & 5.85\\\hline
Genome 8& 99613856  & 	5.74	&87102639   &6.41	&87102639&	6.30 \\\hline
Genome 9& 71876739  &	4.15    &62998800  &4.58 &	62998800 8&	5.50 \\\hline
Genome 10& 75324432 & 	4.36	&66084958  &4.81	&66084958&	4.73 \\\hline
Genome 11& 80766360 &	4.66    &70620666  &5.41&	70620666&	5.04 \\\hline
Genome 12& 74560825 &	4.31    &65359604  &4.75	&65359604&	4.67\\\hline
Genome 13& 71562941 & 	4.14	&62758225  &4.56	&	62758225 &	4.50\\\hline
Genome 14& 78261299 & 	4.51    &68354090  &4.8&	68354090& 4.89  \\\hline
Genome 15& 324439242& 	18.77    &286008420 & 20.66	 &	288866525 &	19.94 \\\hline
Genome 16& 703734840&  40.77    &618859291 & 45.08  &	625042576 &	43.18 \\\hline
Genome 17& 901032667&  52.08    &791840455 &	57.36 &	799757406 &	55.12 \\\hline
Genome 18& 983434816&  56.98    &867299889 &	62.42 &	875969422 & 60.70 \\\hline

\bottomrule
\end{tabular}
%\end{center}
\end{table}

\begin{table}[tpbh]
\small

>>>>

\documentclass[preprint,12pt]{elsarticle}%preprint,

@@

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algcompatible}
\usepackage{multirow}

@@

%\author{}
\author[a,b]{Youcef Gheraibia\corref{cor1}}
\ead{Y.Gheraibia@hull.ac.uk}
\author[a]{Sohag Kabir\corref{cor2}}
\ead{s.kabir@hull.ac.uk}
\author[c]{Mohammad Kaykobad}
\author[d]{Abdelouahab Moussaoui}
\author[b]{Smaine Mazouzi}

\address[a]{Department of Computer Science, University of Hull, Hull HU6 7RX, UK}

\address[b]{Department of Computer Science, Badji Mokhtar-Annaba University, P.O. Box 12, 23000 Annaba, Algeria}

\address[c]{Department of Computer Science and Engineering, Bangladesh University of Engineering and Technology, Dhaka 1000, Bangladesh}

\address[d]{LRIA laboratory, Department of Computer Science, University of Setif, 19000, Algeria}

@@

\begin{keyword}
Data Compression, Huffman code, Cost Considering Approach, Genetic algorithm, Biological data.
%% keywords here, in the form: keyword \sep keyword

@@

\section{Background}
\label{sec2}
\subsection{Issues in Biological Data Transmission}
The size of biological data including DNA sequences increase with an ever expanding rate and will be bigger and bigger in the future. These biological data are stored in biology databases. The exponential growth of these databases become a big problem to all biological data processing methods~\cite{Doug08}.
Different operations are applied to these data such as searching ~\cite{val10},e-mail attachment~\cite{chr09}, alignment ~\cite{che03}, and transmission on distributed computing platform~\cite{cha14}. Interestingly, biological data compression can play a key role in all sort of biological data processing.

@@

Applications of Huffman code are pervasive throughout
computer science. The algorithm to completely perform Huffman encoding and decoding is explained in \cite{Amst86}. It can be used effectively where there is a need for a compact code to represent a long series of a relatively small number of distinct bytes. For example, Table \ref{table2} shows 8 different ASCII characters, their frequencies, ASCII codes and the codewords generated for those symbols using Huffman code. It is seen from the table that the codeword to represent each character is compressed and the most frequent character gets the shortest code. In this example, the compression ratio obtained by Huffman code is 64.16\%.

@@

\section{The Proposed Approach}
\label{sec3}
%\subsection{}

@@

\subsection{Cost Considering Approach of Huffman code}
The classical Huffman algorithm aims at reducing total number of bits and it constructs a tree in a bottom up fashion. It is shown in \cite{golin98} that if the costs of letters are considered unequal then the straightforward bottom up greedy approach does not work. Authors in \cite{Kab14} uses a top down approach to build a binary tree considering unequal letter cost of bits. They considered cost (length) of 0 and 1 as integer constants $\alpha$ and $\beta$,   $\alpha < \beta$. Using the analogy of Morse code's `$\cdot$' and `$-$', the value of $\alpha$ and $\beta$ is set as 1 and 3 respectively. The complete algorithm to obtain an optimal prefix-free code for unequal letter cost is shown below (algorithm 1). The input to the algorithm are the distinct triplets contained in the genome sequence to be encoded and their frequencies. The process of creating the binary tree starts with a single node (root node) and it is initialised with cost 0.  After that, two child (leaf) nodes are created for the root node, i.e., level of the tree is increased by one. Cost of the left child is calculated as the summation of the cost of its parent node and the length of the left arc, and cost of the right child is calculated as the summation of the cost of its parent node and the length of the right arc. Length of left and right arcs are actually the cost (length) of 0 ($\alpha$) and 1 ($\beta$) respectively. The next step is to take a child node with least cost and create two child nodes for it and make it a parent node. In this way, in every iteration the child node with least cost becomes a parent node with two new child nodes. Creation of new child nodes is stopped when total number of child nodes become equal to the number of distinct triplets needed to be encoded.

\begin{algorithm}[!btph]
\caption{Cost Considering Algorithm (CCA)}
\label{alg1}

@@

\FOR{each distinct symbol $i$} 
\STATE $ Enqueue\left(max\_Q , frequency~[~i~]\right)$
\ENDFOR
\STATE create a root node
\STATE $cost~[~root~]\leftarrow 0$
\STATE $Enqueue\left(min\_Q , cost~[~root~]\right)$
\STATE Define costs of the left and right child of the binary tree
\REPEAT 
\STATE $cost\_of\_parent\_node\leftarrow Dequeue\left(min\_Q\right)$
\STATE create $left$ and $right$ child for this node
\STATE $cost~[~left\_child~]\leftarrow cost\_of\_parent\_node+left\_child\_cost$
\STATE $Enqueue\left(min\_Q , cost~[~left\_child~]\right)$
\STATE $cost~[~right\_child~]\leftarrow cost\_of\_parent\_node+right\_child\_cost$
\STATE $Enqueue\left(min\_Q , cost~[~right\_child~]\right)$
\STATE Mark parent node as explored
\UNTIL{$2\left(n-1\right)$ nodes are created}
\WHILE{$min\_Q\neq\emptyset$} 
\STATE $leaf\_node\leftarrow Dequeue\left(min\_Q\right)$
\STATE $frequency[leaf\_node]\leftarrow Dequeue\left(max\_Q\right)$
\ENDWHILE
\FOR{each parent node $j$} 
\STATE $frequency~[~j~]\leftarrow frequency~[~left\_child~]+frequency~[~right\_child~]$
\ENDFOR
\REPEAT
\IF{conflict between nodes} 
\STATE resolve conflict by swapping conflicted nodes
\STATE calculate and reassign cost of all affected nodes
\STATE calculate and reassign frequency of all affected nodes 
\ENDIF
\UNTIL{all conflicts are resolved }
%\algstore{myalg}
\end{algorithmic}
\end{algorithm}

Now the tree $T$ is constructed and the cost of the tree actually depends on how the frequencies are assigned to the leaf nodes. The overall cost will be minimised if the leaves with highest cost always have smaller or equal weight (frequency). To fulfil this condition the leaves of the $T$ are enumerated in non-decreasing order of their cost, i.e., $cost(l_1)\leq cost(l_2)\leq \ldots \leq cost(l_n),$ and that $f_1 \geq f_2 \geq \ldots\geq f_n$, where $l_i$ and $f_i$ are leaf nodes and frequencies of distinct triplets respectively $for~ i=1,2,\ldots n$ . The frequency or weight of parent nodes are calculated as the sum of the frequencies of its child nodes, and it continues upwards until the root node is reached. After that, the algorithm checks for any possible conflicts between all pair of nodes. Two nodes are considered as conflicted if the node with a higher cost has higher frequency violating the above condition, i.e., if $cost(l_i)> cost(l_j)~and~f(l_i)>f(l_j)$, then there remains a conflict.  If there remains a conflict between nodes, then it is resolved by swapping the nodes and recalculating the cost of the tree downward and frequency of the nodes upward.  When all the conflicts if existed are resolved then the algorithm generates codes for each of the distinct triplets.

\subsection{Optimisation of the Codes}
\subsubsection{Problem formulation}
The problem of finding the best allocation of codes to each triplet can be   modelled as an assignment problems, the problem is formulated as follows :\\\\
\textbf{Definition:} Given a set of codes $C=\left\{C_{1},C_{2}...C_{n}\right\}$, and a set of frequencies $F=\left\{F_{1},F_{2}...F_{n}\right\}$. For each code we have the length of the code $|C_{i}|$ (number of bits) and the cost of the code $S_{C_{i}}$, the objective is to assign to each frequency a code in order to minimise the total number of bits, while respecting the initially assigned total cost $S_{t}$ with a given penalty $\lambda\rightarrow [0,1]$. This penalty coefficient represents the allowed amount of cost that can be sacrificed to optimise the total number of bits.    
\\
The Objective Function is to:

@@

\subsubsection{Basic Genetic Algorithm}
Genetic Algorithm (GA) is a bio-inspired meta-heuristic algorithm developed by \cite{gen1}. GA is a stochastic optimisation algorithm imitate the natural evolution process of genomes. GA started by generating an initial population of random feasible solutions. The optimisation process of GA takes the initial population and generates a new population based on it.  The process can be described as follow:\\ 
Firstly, select two or more solutions from the current population by using one of the well-known selection techniques \cite{bli95}. These selected solutions will be considered as  parents. The second operation of the genetic algorithm is the crossover, which takes these parents as inputs to generate other new solutions considered as children. The third operation of the genetic algorithm is the mutation which ensures a good diversification in the search process. The new solutions can be mutated according to a given mutation probability. The mutation operation changes the value of one or more positions in the solution. The quality of each solution is verified by the fitness function which controls the evolution of the GA population by deletion of the worst solutions and insertion of the good solutions among  parents and sons. This processes is repeated until the stopped criteria is reached which can be the number of generation or if the population is stabilised.

\subsubsection{Optimised cost considering algorithm}
The main objective of the optimised cost considering algorithm (OCCA)is to apply GA to find an optimal allocation of the codewords to each triplets to reduce the total number of bits. In GA, the encoding of the solution into a chromosome is one of the most important issues in obtaining good optimisation results. The OCCA uses two fixed length arrays of size ``64" which is the number of combination for all nucleotides. The first array contains the frequencies of each triplets and the second array contains the cost of the codewords assigned to each triplets. Our genetic algorithm utilises the two arrays and uses the index of each entries in the allocation process.

The genetic algorithm is a stochastic algorithm based on random evolution process. The algorithm 2 describes the outline of the OCCA.   Generally the initial population is generated in a random affectation of codes to different triplets. In the OCCA algorithm, the population contain firstly the affectation given by the CCA. The rest of the population are randomly generated, all these generated solutions must satisfy the initial cost constraints, which is the cost ($S_t$) of the solution produced by CCA  and previously added to the population with a give penalty (step 1).

The evolution of the population is the key of the genetic optimisation algorithm. During each generation the process start with the selection of a part (set of solutions) of the population to bread a new generation. In the literature many selection methods have been proposed to guide the population evolution \cite{bli95}. The different methods for selecting a part (set of solutions) of the population vary from a random selection method to heuristic based selection method. We have chosen to select randomly the part of the population to be processed as the heuristic methods are very time-consuming (step 3).
After that the operations of genetic algorithm is applied on the initial population to generate a new generation of the population (see Fig.2). Firstly the crossover operation is applied to these two selected solutions (considered as parents) to generate two new solutions (considered as children)(step 4). In the literature many crossover techniques have been used in genetic algorithm \cite{osa14}, such as one-point crossover which divides the chromosomes into two fragments and recombines the second fragments with the other chromosome's second fragment. The two-point crossover divides the chromosomes into three fragments and recombines the middle fragment with  the middle fragment of the other chromosome. There are many other crossover techniques to allow a good convergence of the algorithm. In our case, we have used the two point crossover with two parameters, the first parameter $p_{c}$ is a random value in [0,63], which represents the first cut point and the second parameter $p_{m}$ is also a random value in [0,63], which represents the number of positions to be crossed. These two random parameters are to ensure a good diversification on the whole search space (step 4).

@@

\begin{figure}[!thpb]
\begin{center}
\includegraphics[scale=0.55]{Images/Drawing_1.jpg}
\caption{Population update for genetic algorithm}
\end{center}
\label{Fig2}
\end{figure}

\begin{algorithm}[!thpb]
\caption{Optimised Cost Considering Algorithm (OCCA)}
\begin{algorithmic}[1]
\State Population initialization (P).
\REPEAT  
\State Select two solutions $S_{1},S_{2}$ form P.
\State Crossover $S_{1},S_{2}$ to generate $S_{11},S_{21}$ (Children).
\State Mutate $S_{11},S_{21}$.
\State Validate children with cost constraint (equation 4). 
\State Add children to population
\State Rank the population by fitness
\State Remove worst candidates until population limit
\UNTIL{Max number of generation not achieved} 
\EndWhile
\State Display the best solution from the population P;
\end{algorithmic}
\end{algorithm}

After the crossover operation, the newly generated children may contain conflict, for example, a single code is allocated to two different frequencies in the solution. To overcome such a conflict, a regulation operation is performed to refine the solution to ensure the correctness of the solution (see Fig.2).
 Secondly these two new solutions are mutated according to a predefined probability $\gamma$, the best value of the mutation rate is very problem specific (step 5). In our case, the value of $\gamma$ is fixed to $0.2$ to explorer a few positions in the solution. The mutation operator is used to maintain genetic diversity from one generation of the population of genetic algorithm to the next. In our case, we have used the mutation as a random swap mutation operator (see Fig.3).  Each newly generated solution must satisfy the cost constraint. The next step is to add these two new solutions (children) to the current population (step 7) (see Fig. 3). Finally, the new population are ranked by fitness (step 8), and the worst solutions are deleted until the initial size of the population is obtained  (step 9). The whole processes are repeated until the maximum number of operation is performed (step 10).

\section{Results And Discussion}
\label{sec4}
The effectiveness of the approach has been evaluated with different real genomic biological data, these genomes were downloaded from a recent version of The National Center for Biotechnology Information (NCBI) available on $(http://www.ncbi.nlm.nih.gov)$ \cite{pruitt2009ncbi}. We focused on the sequences alone, ignoring any header and any other exogenous information. In table 3, the different data sets are described with the size  of each of them in megabytes (MB) and the references on the biological data bank.

\begin{table}[!thpb]

\small
\label{table3}
\caption{Datasets description}
%\begin{center}
\centering
\begin{tabular}{c  c  c c}
\toprule
$\textbf{Data sets}$ &$\textbf{Name}$ &	$\textbf{Size (MB)}$ &	$\textbf{Reference}$ \\\hline
Genome 1& Mycobacterium smegmatis &  6.66 & CP009496  \\\hline

@@

\multirow{2}{*}{Genome 16}& Danio rerio & \multirow{2}{*}{56.14} &\multirow{2}{*}{CM002885}\\ & chromosome 1 &\\
\hline
\multirow{2}{*}{Genome 17}&Homo sapiens & \multirow{2}{*}{76.64 } &\multirow{2}{*}{  CM000680  }\\    & chromosome 18 &\\
\hline
\multirow{2}{*}{Genome 18}&Homo sapiens & \multirow{2}{*}{99.94} &\multirow{2}{*}{CM000684   }\\ & chromosome 22&\\
\hline
%\bottomrule
\end{tabular}
%\end{center}
\end{table}

\begin{table}[h]
\renewcommand{\arraystretch}{1.1}
\small
\label{table4}
\caption{Comparison of performance among classical Huffman code, CCA, and OCCA without penalty}
%\begin{center}

\begin{tabular}{c  c c  c c  c c}
\hline
 & \multicolumn{2}{c}{Huffman Algorithm} & \multicolumn{2}{c}{CCA }& \multicolumn{2}{c}{OCCA \textit{($\lambda$=0)}}\\\hline
$\textbf{Data sets}$ & $\textbf{Cost}$	& $\textbf{Size(MB)}$ &	$\textbf{Cost}$&	$\textbf{Size(MB)}$&$\textbf{Cost}$&	$\textbf{Size(MB)}$
\\\hline
Genome 1& 76787151  &	4.44	&67416213   & 4.92	& 67416213 & 4.82 \\\hline
Genome 2& 100425402 &  	5.81	& 88430665	& 6.51	& 88430665&  6.35 \\\hline
Genome 3& 75940155  &	4.39    &66745619	& 4.87&	66745619 & 4.79 \\\hline
Genome 4& 103552729 &	5.96    &90821835&	6.69 &	90821835  & 6.47\\\hline
Genome 5& 82234926  &	4.74	&71963876	& 5.27 &	71963876 &5.17 \\\hline
Genome 6& 83454842  &   4.81	&73038795	& 5.34	&73038795 & 5.24\\\hline
Genome 7& 92539488  &	5.36    &81416359   &5.93	&81416359 & 5.85\\\hline
Genome 8& 99613856  & 	5.74	&87102639   &6.41	&87102639&	6.30 \\\hline
Genome 9& 71876739  &	4.15    &62998800  &4.58 &	62998800 &	4.49 \\\hline
Genome 10& 75324432 & 	4.36	&66084958  &4.81	&66084958&	4.73 \\\hline
Genome 11& 80766360 &	4.66    &70620666  &5.41&	70620666&	5.04 \\\hline
Genome 12& 74560825 &	4.31    &65359604  &4.75	&65359604&	4.67\\\hline
Genome 13& 71562941 & 	4.14	&62758225  &4.56	&	62758225 &	4.50\\\hline
Genome 14& 78261299 & 	4.51    &68354090  &4.8&	68354090& 4.89  \\\hline
Genome 15& 324439242& 	18.77    &286008420 & 20.66	 &	286008420 &	19.94 \\\hline
Genome 16& 703734840&  40.77    &618859291 & 45.08  &	618859291 &	43.18 \\\hline
Genome 17& 901032667&  52.08    &791840455 &	57.36 &	791840455 &	55.12 \\\hline
Genome 18& 983434816&  56.98    &867299889 &	62.42 &	867299889 & 60.70 \\\hline

%\bottomrule
\end{tabular}
%\end{center}
\end{table}

\begin{figure}[h]
\begin{center}
\includegraphics[scale=0.45]{Images/drawing4_1.jpg}
\caption{Convergence of OCCA for \textit{Genome 2}}
\end{center}
\label{Fig5}
\end{figure}

\begin{table}[tpbh]
\small
\renewcommand{\arraystretch}{1.1}
\caption{Effects of  the penalty on the compression performance of the OCCA}
%\begin{center}
\centering
\begin{tabular}{c c c c c }
\toprule
$\textbf{Data sets}$ & $\textbf{Cost}$	& $\textbf{Size(MB)}$ &		\textbf{$\lambda (\%)$} &		\textbf{Compression ratio $\%$} \\\hline
Genome 1 &	 70760174 &4.50& 4\% & 32.72\%\\\hline
Genome 2 &	92010490& 5.91 & 5\%& 28.79\%\\\hline
Genome 3&	69421783& 4.47 &3\%& 26.84\%\\\hline
Genome 4&	96294638& 5.99 & 5\%& 29.85\%\\\hline
Genome 5&	74855668 &4.81& 5\%& 27.77\%\\\hline
Genome 6&	76738330& 4.86  &4\%& 28.10\%\\\hline
Genome 7&	84667949 &5.47& 3\%& 26.28\%\\\hline
Genome 8&	92416243 &5.76& 5\%& 30.35\%\\\hline
Genome 9&	66145783&4.21 & 4\%& 26.65\%\\\hline
Genome 10&	68751359& 4.44 &3\%& 26.63\%\\\hline
Genome 11&	72737300 &4.79& 2\%&26.08\%\\\hline
Genome 12&	67981988 & 4.38&3\%& 26.63\%\\\hline
Genome 13&	65896779& 4.18 &4\%& 27.30\%\\\hline
Genome 14&	71762305 &4.55 &4\%&27.54\% \\\hline
Genome 15&297188002 &19.31 & 3\%& 24.51\%\\\hline
Genome 16&643785655 & 41.68 & 3\%& 25.75\% \\\hline
Genome 17&823506882 & 53.05 & 3\%& 30.78\%\\\hline
Genome 18&901515198 &58.99 & 3\%& 40.97\%\\
\bottomrule
\end{tabular}
%\end{center}
\label{table5}
\end{table}
